### 2021-08-04 msc dissertation code.Rmd
latest

---
title: "msc dissertation"
author: "Halleyson Li"
date: "'r Sys.Date()'"
output: html_document
line-citations: apalike
linkcolor: blue
frontsize: 12pt
---

\newpage

# *Notes*{-}
1. The GitHub repository for this project could access through [here](https://github.com/halleyson/msc-dissertation.git)

# Loading data

```{r, message=FALSE}

library(readr)

pennycook1 <- read_csv("Pennycook et al._Study 1.csv")
pennycook2 <- read_csv("Pennycook et al._Study 2.csv")
# Quick check
head(pennycook2)

```

# General steps to regenerate headline result (study 2)
## data cleaning

```{r, message=FALSE}

# rename columns that contains special characters
colnames(pennycook2)
names(pennycook2)[1] <- "condition"

#conclude a sheet with only the relevant information 
#condition 1 = control; condition 2 = treatment

#Choose with the column position 
library("dplyr")
pennystudy2r1 <- pennycook2 %>% 
  select (c(1, 2, 3, 4, 57, 60, 250, 252, 284))

#re-code sharing intentions (fake/real) using the scale midpoint (1-3 = 0, 4-6 = 1)
pennystudy2r1$Fake[pennystudy2r1$Fake > 0 & pennystudy2r1$Fake < 4] <- 0
pennystudy2r1$Fake[pennystudy2r1$Fake > 3 & pennystudy2r1$Fake < 7] <- 1
pennystudy2r1$Fake

pennystudy2r1$Real[pennystudy2r1$Real > 0 & pennystudy2r1$Real < 4] <- 0
pennystudy2r1$Real[pennystudy2r1$Real > 3 & pennystudy2r1$Real < 7] <- 1
pennystudy2r1$Real

#create veracity column (Real - Fake)
pennystudy2r1['veracity']=pennystudy2r1['Real'] - pennystudy2r1['Fake']

#successful screen1 test (result = 2)
pennystudy2r1['screen1']=pennystudy2r1['screen1_4'] + pennystudy2r1['screen1_7']

#successful screen2 test (result = 2)
pennystudy2r1['screen2']=pennystudy2r1['screen2_3'] + pennystudy2r1['screen2_5']

#successful screen3 test (result = 3)
pennystudy2r1['screen3'] = na_if(pennystudy2r1$screen3_2,1)
pennystudy2r1['screen3'] = na_if(pennystudy2r1$screen3,2)
pennystudy2r1['screen3'] = na_if(pennystudy2r1$screen3,4)
pennystudy2r1['screen3'] = na_if(pennystudy2r1$screen3,5)
#check if screen 3 successful
mean(pennystudy2r1$screen3, na.rm = TRUE)

#re-code screen test: replace 2 and 3 to 1
pennystudy2r1["screen1"][pennystudy2r1["screen1"] == 2] <- 1
pennystudy2r1["screen2"][pennystudy2r1["screen2"] == 2] <- 1
pennystudy2r1["screen3"][pennystudy2r1["screen3"] == 3] <- 1

#re-code condition control = 0 treatment = 1
pennystudy2r1["condition"][pennystudy2r1["condition"] == 1] <- 0
pennystudy2r1["condition"][pennystudy2r1["condition"] == 2] <- 1

#replace NA to 0 in screen 1,2,3

pennystudy2r1[c("screen1","screen2","screen3")][is.na(pennystudy2r1[c("screen1","screen2","screen3")])] <- 0

#create dummy column, add screen 1,2,3 together. Screen = 3 indicate success in all attention check
pennystudy2r1['dummycheck']=pennystudy2r1['screen1'] + pennystudy2r1['screen2'] + pennystudy2r1['screen3']

#like the researcher did, change data frame from wide to long
##first drop unnecessary column
pennys2r1long <- subset(pennystudy2r1, select = c(1,2,3,4,10,14))

##Add ID column
pennys2r1long <- tibble::rowid_to_column(pennys2r1long, "ID")

## change from wide to long
library(tidyr)

keycol <- "newstype"
valuecol <- "measurement"
gathercols <- c("Fake","Real")

pennys2r1long <- gather_(pennys2r1long, keycol, valuecol, gathercols)

##no idea why it doesn't show groups (control vs treatment) so I merge original condition to the new one

merge(pennystudy2r1, pennys2r1long, by.x='condition', by.y='ID')

##organize data by ID, then reorder the column
pennys2r1long <- pennys2r1long[order(pennys2r1long$ID),]

pennys2r1long <- pennys2r1long[, c(1, 2, 6, 7, 4, 5, 3)]

##separating the data based on number of successful attention check

### attention check above and equal to 1
pennys2r1longa1 <- subset(pennys2r1long, dummycheck %in% c("1","2","3"))

### attention check above and equal to 2
pennys2r1longa2 <- subset(pennys2r1long, dummycheck %in% c("2","3"))

### attention check equal to 3
pennys2r1longa3 <- subset(pennys2r1long, dummycheck %in% c("3"))

```

## reproduction of results (interaction between headline veracity and treatment)
### without considering attention check
```{r, message=FALSE}
install.packages("sandwich") 
install.packages("stargazer")
install.packages("lmtest")
library(lmtest)
library(stargazer) 
library(sandwich)

lm1 <- lm(veracity ~ newstype + newstype*condition, data = pennys2r1longa2)
summary(lm1)

stargazer(lm1, type='text')

coeftest(lm1, vcovHC)

#z scored
scale(pennystudy2r1$Discern)

#calculate the effect size cohen's d
##since they need to compare true vs false headline in different condition, thus, create a new data frame that only contain treatment and control with the attention check above and equal to 2, as stated in the research

install.packages("effsize")
library(effsize)

pennys2r1a2cohen <- subset(pennystudy2r1, dummycheck %in% c("2","3"))
pennys2r1a2cohenc <- subset(pennys2r1a2cohen, condition %in% c("0"))
pennys2r1a2cohent <- subset(pennys2r1a2cohen, condition %in% c("1"))

cohen.d(pennys2r1a2cohenc$Fake, pennys2r1a2cohenc$Real, na.rm = TRUE)
cohen.d(pennys2r1a2cohent$Fake, pennys2r1a2cohent$Real, na.rm = TRUE)

#condition vs measurement

pennys2r1a2cohenc <- subset(pennys2r1longa2, condition %in% c("0"))
pennys2r1a2cohent <- subset(pennys2r1longa2, condition %in% c("1"))

cohen.d(pennys2r1a2cohenc$condition, pennys2r1a2cohenc$measurement, na.rm = TRUE)
cohen.d(pennys2r1a2cohent$condition, pennys2r1a2cohent$measurement, na.rm = TRUE)

```

### reproduction of figure 2 without considering attention check

### graph 1: including all
```{r, message=FALSE}

library(ggplot2)

#Since the plot not specify which condition they use, then I need to plot all 4 graph in different condition

#condition 1: all

pennystudy2plot1 <- pennys2r1long %>%
  select (c(1, 2, 3, 4))

#Find out the sum for each type of news in different condition

p1 <- aggregate(measurement~condition+newstype, data = pennystudy2plot1, na.rm = TRUE, FUN = sum)
p1 <- transform(p1, percent = ave(measurement, condition, FUN = function(x) paste0(round(x/428, 4)*100)))

##double check for correctness
sum(pennystudy2plot1$measurement,na.rm = TRUE)

#since the obs is 1712, thus 1712/2 = 856 participants. So to get percentage of sharing intention each fake and real news need to divided by 856/2 = 428
ggplot(p1, aes(x = condition, y = as.numeric(percent), fill = newstype)) + 
  geom_bar(stat = 'identity',position = 'dodge') + 
  geom_text(
    aes(label = as.numeric (percent)), vjust = 1.5, position = position_dodge(0.9))
```
### graph 2 attention check above and equal to 1

```{r, message=FALSE}

#condition 2: above and equal to 1

pennystudy2plot2 <- pennys2r1longa1 %>%
  select (c(1, 2, 3, 4))

#separating the sheet

pennystudy2plot2c <- subset(pennystudy2plot2, condition %in% c("0"))
pennystudy2plot2t <- subset(pennystudy2plot2, condition %in% c("1"))

#Find out the sum for each type of news in different condition
p2 <- aggregate(measurement~condition+newstype, data = pennystudy2plot2, FUN = sum)

##double check for correctness
sum(pennystudy2plot2$measurement,na.rm = TRUE)

#since the obs is 1664, thus 1664/2 = 832 participants. So to get percentage of sharing intention each fake and real news need to divided by 832/2 = 416
ggplot(p2, aes(x = condition, y = measurement/428, fill = newstype)) + 
  geom_bar(stat = 'identity',position = 'dodge') + 
  geom_text(
    aes(label = round(measurement/428, digits = 3)), vjust = 1.5, position = position_dodge(0.9))+
    scale_y_continuous(labels = scales::percent)
```

### graph 3 attention check above and equal to 2
```{r, message=FALSE}
#condition 3: above and equal to 2
pennystudy2plot3 <- pennys2r1longa2 %>%
  select (c(1, 2, 3, 4))

#Find out the sum for each type of news in different condition

p3 <- aggregate(measurement~condition+newstype, data = pennystudy2plot3, na.rm = TRUE, FUN = sum)
p3 <- transform(p3, percent = ave(measurement, condition, FUN = function(x) paste0(round(x/428, 4)*100)))

##double check for correctness
sum(pennystudy2plot3$measurement,na.rm = TRUE)

#plot the graph under the proportion of overall observation (856)
ggplot(p3, aes(x = condition, y = as.numeric(percent), fill = newstype)) + 
  geom_bar(stat = 'identity',position = 'dodge') + 
  geom_text(
    aes(label = as.numeric (percent)), vjust = 1.5, position = position_dodge(0.9))
```
### graph 3 attention check above and equal to 2: mean
```{r, message=FALSE}
#condition 3: above and equal to 2
pennystudy2plot3 <- pennys2r1longa2 %>%
  select (c(1, 2, 3, 4))

pennystudy2plot3c <- subset(pennystudy2plot3, condition %in% c("0"))
pennystudy2plot3t <- subset(pennystudy2plot3, condition %in% c("1"))

#Find out the sum and percentage for each type of news in each condition

##control, percentage = 658/2 = 329
p31 <- aggregate(measurement~condition+newstype, data = pennystudy2plot3c, na.rm = TRUE, FUN = sum)
p31 <- transform(p31, percent = ave(measurement, condition, FUN = function(x) paste0(round(x/329, 4)*100)))

##treatment, percentage = 634/2 = 317
p32 <- aggregate(measurement~condition+newstype, data = pennystudy2plot3t, na.rm = TRUE, FUN = sum)
p32 <- transform(p32, percent = ave(measurement, condition, FUN = function(x) paste0(round(x/317, 4)*100)))

#merge two table together under same column name
p3 <- rbind(p31,p32)

##double check for correctness
sum(pennystudy2plot3$measurement,na.rm = TRUE)

#plot the graph
ggplot(p3, aes(x = condition, y = as.numeric(percent), fill = newstype)) + 
  geom_bar(stat = 'identity',position = 'dodge') + 
  geom_text(
    aes(label = as.numeric (percent)), vjust = 1.5, position = position_dodge(0.9))
```

### graph 4 attention check equal to 3
```{r, message=FALSE}
#condition 3: equal to 3
pennystudy2plot4 <- pennys2r1longa3 %>%
  select (c(1, 2, 3, 4))

#Find out the sum for each type of news in different condition
p4 <- aggregate(measurement~condition+newstype, data = pennystudy2plot4, FUN = sum)

##double check
sum(pennystudy2plot4$measurement,na.rm = TRUE)

#since the obs is 460 for attention = 3, thus 460/2 = 230 participants. So to get percentage of sharing intention each fake and real news need to divided by 230/2 = 115
ggplot(p4, aes(x = condition, y = measurement/115, fill = newstype)) + 
  geom_bar(stat = 'identity',position = 'dodge') + 
  geom_text(
    aes(label = round(measurement/115, digits = 3)), vjust = 1.5, position = position_dodge(0.9))+
    scale_y_continuous(labels = scales::percent)
```